Appendix

### 2024-09-16 manuscript/manuscript.Rmd
Added note normalisation prior to transformation.

Location- and scale-invariant transformations are realised by simultaneously optimising three parameters, i.e. transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$.
Alternatively, standardisation of a numeric variable (e.g., through subtracting its median value and division by its interquartile range) prior to conventional power transformations may achieve a similar effect in reducing sensitivity to the distribution's location and scale.

@@

As a limitation, we did observe several numerical issues for optimisation criteria other than MLE (appendix B).

>>>>

Location- and scale-invariant transformations are realised by simultaneously optimising three parameters, i.e. transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$.
Alternatively, standardisation of a numeric variable (e.g., through subtracting its median value and division by its interquartile range) prior to conventional power transformations may achieve a similar effect in reducing sensitivity to the distribution's location and scale.
While this alternative helps prevent numerical issues -- provided that normalisation does not lead to negative values for Box-Cox transformation -- location- and scale-invariant transformations seem to provide an overall better transformation to normality (Appendix F).

@@

As a limitation, we did observe several numerical issues for optimisation criteria other than MLE (Appendix B).

### 2024-09-16 manuscript/manuscript_appendix.Rmd
Added note normalisation prior to transformation.

```{r include = FALSE, eval = TRUE}

data <- .assess_standardisation_before_transformation()
```

>>>>

An alternative to location- and scale-invariant transformations is normalising feature distributions prior to conventional transformations.
Table \ref{tab:normalisation-before-transformation-appendix} shows residual errors, after transformation to normality, of the five features from real-world datasets presented previously in Appendix D and the main manuscript.
In these examples location- and scale-invariant transformations have similar or lower residual errors compared to errors resulting from normalisation prior to transformation.

```{r include = FALSE, eval = FALSE}

data <- .assess_standardisation_before_transformation()
data <- data[transformation %in% c("none", "conventional", "conventional (z-score norm.)", "conventional (robust scaling)", "invariant"), mget(c("residuals", "transformation", "dataset"))]
data[, "residuals" := round(residuals, digits = 1L)]
data.table::dcast(data, dataset ~ transformation, value.var = "residuals")
```
```{=latex}
\begin{table}
\begin{center}
\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. conv.: conventional; norm.: normalisation; rob.: robust}
\label{tab:normalisation-before-transformation-appendix}
\begin{tabular}{l | c c c c c}

\toprule
feature & none & conventional & conv. (z-score norm.) & conv. (rob. scaling) & invariant \\

\midrule
age                     &  16.5 &  11.5 &  11.5 &  11.3 &   8.8 \\
penguin body mass       &  48.0 &  32.2 &  33.3 &  32.2 &  26.8 \\
latitude                & 328.1 & 319.0 & 326.2 & 324.5 & 326.4 \\
fuel efficiency         &  54.5 &  55.3 &  49.0 &  53.3 &  44.0 \\
arterial wall thickness & 110.1 &  30.0 &  19.3 &  31.8 &  12.2 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

# References
